The design and implementation of standards-based Grid single sign-on using federated identity

### 2009-04-30 doc/sec/papers/GRID09/paper.tex
a little change to the paper

The rest of this paper is organized as follows: Section~\ref{sec:arcmiddleware} presents the 
ARC Grid middleware, and in particular its new architecture. Section~\ref{sec:intergrationSAML2SSO} 
describes the solution of using federated identity for 
Grid authentication by using \texttt{SAML2SSO} profile; Section~\ref{sec:fedidtoX509} describes
how to obtain X.509 credentials based on federated identity, and how to use such for 
accessing multiple Grid resources; Section~\ref{sec:perfeval} is dedicated to performance evaluation;
Section~\ref{sec:relatedwork} describes the related work; Section~\ref{sec:conclusion} 
contains conclusions and future work outlook.

@@

The key part of the new approach is a lightweight Web 
Service container called Hosting Environment Daemon (HED), which provides hosting environment 
for various services at application level, as well as a number of modules to support flexible, 
interoperable, and efficient communication mechanism for building SOAP-based Web Services. 
The design of the HED is built around the idea of flexibility and modularity,
such that the application service developers can simply concentrate on the application 
level Web Service implementation by only using the core minimum amount of components. It 
also simplifies work on the middleware level: for example, it makes possible to implement another 
communication protocol or authentication mechanism. A system administrator can easily 
configure and deploy the middleware and application for different kinds of requirements 
without having to know much about the implementation.

@@

\begin{figure}
\includegraphics[width=1.0\columnwidth]{SAML2SSO_UML.pdf}
\caption{Sequence diagram of SAML2.0 SSO profile integration in SOAP invocation 
between ARC WS Client and Web Service}
\label{fig:SAML2SSOUML}
\end{figure}

Figure~\ref{fig:SAML2SSOUML} shows how the SAML2.0 SSO profile is integrated
into the SOAP invocation between ARC WS Client and Web Service. Steps 1 to 5 in
Figure~\ref{fig:SAML2SSOUML} are similar to those steps depicted in the Web Browser SSO profile, with
the difference that we specify the way how does the service provider determine which identity
provider to use (identity provider discovery) by enabling the client to send the IdP name to the service
provider. We also specify the way how does the client (user agent)
authenticate against a service provider by implementing the form-based HTTP authentication which uses username and
password for authentication between the user agent and a service provider (rather than the X.509 client authentication),
which is compatible with the Username/Password login handler of the Shibboleth IdP.

@@

There are a several benefits with integrating the SAML2.0 SSO profile for
Grid/Web service authentication. Firstly, the client certificate authentication can be switched off, so that
users would not need to apply and maintain X.509 credentials. Secondly, some implementations of
identity provider, such as Shibboleth IdP, can cache the authentication result through session management
once the user has succeeded to authenticate, and for a short period this authentication result
is valid, so that the next time the user authenticates against IdP, the user agent can just feed
IdP with the last successfully authenticated session's identity rather than feed IdP with
the username and password again. So the user (or the client on behalf of that user) can travel across multiple
security domains by only providing his name and password once, which is also a characteristic of
single sign-on. Lastly, even though there are several implementations of identity providers,
most of them are standard-compliant, so the solution implemented in ARC middleware can
interoperate with other identity provider implementations with minimum change.

@@

\subsection{Short-lived credential service}
\label{sec:slcs}
The sequence diagram of a SLCS process is the same as the diagram shown in 
Figure~\ref{fig:SAML2SSOUML}, with the SOAP request (the message of step 8 of 
Figure~\ref{fig:SAML2SSOUML}) including an X.509 request, and the SOAP response (the message of step 10 
of Figure~\ref{fig:SAML2SSOUML}) including a signed X.509 certificate and a CA certificate that is used 
to sign this X.509 certificate. In detail, the SLCS client firstly accomplishes the SAML2SSO profile, 
then sends the SOAP request to the SLCS service with the X.509 request embedded. The SLCS
service makes access control decision according to the SAML assertion stored in the session
context, and issues a certificate (with 12 hours lifetime) with SAML attributes as the X.509
certificate extension. The SLCS client then gets the SOAP response with the X.509
certificate embedded.

@@

\begin{figure}
\includegraphics[width=1.0\columnwidth]{Delegation_UML.pdf}
\caption{Sequence diagram of the usage of delegation service}
\label{fig:DelegUML}
\end{figure}

>>>>

The rest of this paper is organized as follows: Section~\ref{sec:arcmiddleware} presents the 
ARC Grid middleware, and in particular its new architecture. Section~\ref{sec:intergrationSAML2SSO} 
describes the solution of using federated identity for 
Grid authentication by using \texttt{SAML2 Web Browser SSO Profile (SAML2SSO)} profile; 
Section~\ref{sec:fedidtoX509} describes
how to obtain X.509 credentials based on federated identity, and how to use such for 
accessing multiple Grid resources; Section~\ref{sec:perfeval} is dedicated to performance evaluation;
Section~\ref{sec:relatedwork} describes the related work; Section~\ref{sec:conclusion} 
contains conclusions and future work outlook.

@@

The key part of the new approach is a lightweight Web 
Service container called Hosting Environment Daemon (HED), which provides hosting environment 
for various services at application level, as well as a number of modules to support flexible, 
interoperable, and efficient communication mechanism for building SOAP-based Web Services. 
The design of the HED is built around the idea of flexibility and modularity,
such that the application service developers can simply concentrate on the application 
level Web Service implementation by only using the core minimum amount of components. It 
also simplifies work on the middleware level: for example, it makes possible to implement another 
communication protocol or authentication mechanism. Meanwhile, a service administrator can easily 
configure and deploy the middleware and application for different kinds of requirements 
without having to know much about the implementation.

@@

\begin{figure}
\includegraphics[width=1.0\columnwidth]{SAML2SSO_UML.pdf}
\caption{Sequence diagram of the integration of SAML2.0 SSO profile in SOAP invocation 
between ARC WS Client and Web Service}
\label{fig:SAML2SSOUML}
\end{figure}

Figure~\ref{fig:SAML2SSOUML} shows how the SAML2.0 SSO profile is integrated
into the SOAP invocation between ARC WS Client and Web Service. Steps 1 to 5 in
Figure~\ref{fig:SAML2SSOUML} are similar to those steps depicted in the SAML2 Web Browser SSO profile, with
the difference that we specify the way how does the service provider determine which identity
provider to use (identity provider discovery) by enabling the client to send the IdP name to the service
provider. We also specify the way how does the client (user agent)
authenticate against a service provider by implementing the form-based HTTP authentication which uses username and
password for authentication between the user agent and a service provider (rather than the X.509 client authentication),
which is compatible with the Username/Password login handler of the Shibboleth IdP.

@@

There are a several benefits with integrating the SAML2.0 SSO profile for
Grid/Web service authentication. Firstly, the client certificate authentication can be switched off, so that
users would not need to apply and maintain X.509 credentials. Secondly, some implementations of
identity provider, such as Shibboleth IdP, can cache the authentication result through session management
once the user has succeeded to authenticate, and for a short period this authentication result
is valid, so that the next time the user authenticates against IdP, the user agent can just feed
IdP with the last successfully authenticated session's identity rather than feed IdP with
the username and password again. So the user (or the client on behalf of that user) can access multiple
security domains by only providing his name and password once, which is also a characteristic of
single sign-on. Lastly, even though there are several implementations of identity providers,
most of them are standard-compliant, so the solution implemented in ARC middleware can
interoperate with other identity provider implementations with minimum change.

@@

\subsection{Short-lived credential service}
\label{sec:slcs}
The sequence diagram of a SLCS process is the same as the diagram shown in 
Figure~\ref{fig:SAML2SSOUML}, with the SOAP request (the message of step 8 of 
Figure~\ref{fig:SAML2SSOUML}) including an X.509 request, and the SOAP response (the message of step 10 
of Figure~\ref{fig:SAML2SSOUML}) including an X.509 certificate and a CA certificate that is used 
to sign this X.509 certificate. In detail, the SLCS client firstly accomplishes the SAML2SSO profile, 
then sends the SOAP request to the SLCS service with the X.509 request embedded. The SLCS
service makes access control decision according to the SAML assertion stored in the session
context, and issues a certificate (with 12 hours lifetime) with SAML attributes as the X.509
certificate extension. The SLCS client then gets the SOAP response with the X.509
certificate and the CA certificate embedded.

@@

\begin{figure}
\includegraphics[width=1.0\columnwidth]{Delegation_UML.pdf}
\caption{Sequence diagram of the usage of delegation service in SOAP invocation 
between ARC WS Client and Web Service}
\label{fig:DelegUML}
\end{figure}
